---
title: 'Example pipeline: fastq to bam'
author: "Sahil Seth"
date: '`r Sys.Date()`'
output:
  pdf_document:
    highlight: default
    number_sections: yes
    toc: yes
    toc_depth: 4
  html_document:
    theme: flatly
    toc: yes
    toc_depth: 4
---


<!--- output:
  packagedocs::package_docs
  packagedocs::package_docs:
    self_contained: true
    toc: true
--->

# Example: Fastq to merged BAM

You may visit [docs.flowr.space](http://docs.flowr.space/index.html) for more details.

If you face any issues, please feel free to raise a [issue on github](https://github.com/sahilseth/flowr/issues).

## Setup up flowr


Requirements:

- R version > 3.1, preferred 3.2
- install flowr from sahilseth.github.io/drat, provides a 
more recent version than CRAN.

```{r eval=FALSE}
#install.packages("params", repos = "http://cran.rstudio.com")
## for a latest stable version (updated every few days):
install.packages("flowr", repos = "http://sahilseth.github.io/drat")
```


After installation run setup(), this will copy the flowr’s helper script to ~/bin. Please make sure that this folder is in your $PATH variable.

```{r message=FALSE}
library(flowr)
setup()
```

Then we need to test whether we are able to submit jobs to the cluster properly.

```{r, eval=FALSE}
## run a test on the local platform
run(x='sleep_pipe', platform='local', execute=FALSE)
## run a test on the HPCC platform (torque, sge, moab, slurm, lsf)
run(x='sleep_pipe', platform='torque', execute=TRUE)
```

**NOTE**: In case the test is not successful, please follow the [advanced configuration](http://docs.flowr.space/install.html) page for more details on how to solve the issues.




### Fetch and download the pipeline

Next, we will download a pipeline which processes multiple 
[fastq files](https://en.wikipedia.org/wiki/FASTQ_format) of a sample 
into a single aligned and merged [BAM file](http://samtools.github.io/hts-specs/SAMv1.pdf).

```
cd ~/flowr/pipelines
base=https://raw.githubusercontent.com/sahilseth/flowr/devel/inst/pipelines
wget $base/fastq_bam_bwa.R
wget $base/fastq_bam_bwa.conf
wget $base/fastq_bam_bwa.def
```

**Reference Genome**: 

One can download the reference genome including indexes of various alignment tools from [Illumina's iGenomes](https://support.illumina.com/sequencing/sequencing_software/igenome.html) website.

You may skip this step, if you already have the genome fasta and related files.

```
mkdir ~/flowr/genomes; cd ~/flowr/genomes
url=ussd-ftp.illumina.com/Homo_sapiens/NCBI/build37.2/Homo_sapiens_NCBI_build37.2.tar.gz
ftp ftp://igenome:G3nom3s4u@$url
tar -zxvf Homo_sapiens_NCBI_build37.2.tar.gz
```

A typical NGS pipeline consists of many steps, each with several parameters. You can modify `fastq_bam_bwa.conf`, specifying paths to various tools and their default options (samtools, bwa, picard and reference 
genome indexes).


<div class="alert alert-info" role="alert">
**Note**: All parameters of this pipeline are
conveniently specified in a tab-delimited format in the **fastq_bam_bwa.conf** file.
</div>

```
## customize parameters, including paths to samtools, bwa, reference genomes etc.
vi fastq_bam_bwa.conf
```


**Example data**: 

You may skip this step if you already have raw reads for a sample, in fastq format.

```
mkdir ~/flowr/genomes; cd ~/flowr/genomes
## for testing puposes one may download example fastq files:
wget http://omixon-download.s3.amazonaws.com/target_brca_example.zip
unzip target_brca_example.zip
```

### Customize flow definition, describing the computing cluster


Next, we need to customize the resource requirements based on the computing platform.
You may refer to the [flow definition](http://docs.flowr.space/docs.html#flow_definition) format for more details.

```
## customize the resource requirements in flowdef:
- need to change: queue, platform
- may change: walltime, memory, CPUs etc.
vi fastq_bam_bwa.def

## read check flowdef (shell)
flowr as.flowdef x=fastq_bam_bwa.def

## OR from R
as.flowdef(x='fastq_bam_bwa.def')
```

**Read and check flowdef**

\begin{small}

```{r message=FALSE, echo=FALSE}
def = as.flowdef("fastq_bam_bwa.def")
kable(def, format = "latex")
```

\end{small}

A flow definition with default values has already been supplied, briefly,

- Submission Type (**sub_type**): determines, how each step is **submitted** to the cluster. All steps except `merging` may have 
multiple subprocess (each of which can run in parallel). Thus, we spread (`scatter`) them across the cluster.
- Previous Jobs (**prev_jobs**): 
  The two `aln` steps of `bwa` may be run in parallel, and its subsequent `sampe` would wait for both. Spcecifically, in case of 
multiple fastq files $i^{th}$ `sampe` step would wait for $i^{th}$ `aln1` and `aln2` steps.
- **cpu_reserved**: Since `aln` can use multiple cores, we provide it 12 cores, and for rest of the steps, 1 core each.
- **walltime**: Merging may take a little longer, so we give it ample amount of time (12 hours). 
Some computing platforms specify time as `hh:mm:ss` and others prefer `hh:mm`, you may need to check with your system admin.
- **memory**: For simplicity we can assign 16GB (16000kb) of memory to each of these steps 
(may be an overkill, please change as necessary).
- **queue**: We use a generic `medium` queue, since if usually exists; please change as needed.
- **platform**: Finally, specify the platform of your computing cluster (moab, lsf, torque, sge, slurm [alpha])

<div class="alert alert-info" role="alert">
**Tip:** Once we define the flow definition correctly, we may not need to change it any further (one time effort).
</div>






### Submit to cluster 

#### Single step cluster submission


<div class="alert alert-info" role="alert">
**Note**: 
Assuming that the pipeline along with its `.def` and `.conf` files is available in `~/flowr/pipelines`. <br>
Also, `.conf` files should have all the correct paths and 
`.def` file should have resource requirements specified correctly.
</div>


```
## get input fastqs
fqs1=~/flowr/genomes/target_brca_example/brca.example.illumina.0.1.fastq
fqs2=~/flowr/genomes/target_brca_example/brca.example.illumina.0.2.fastq

## submit to the cluster
flowr run x=fastq_bam_bwa fqs1=$fqs1 fqs2=$fqs2 samplename=samp execute=TRUE

## change the platform specified in flowdef
flowr run x=fastq_bam_bwa fqs1=$fqs1 fqs2=$fqs2 samplename=samp execute=TRUE platform=slurm
```

```{r eval=FALSE}
library(flowr)
fqpath = "~/flowr/genomes/target_brca_example"
## demonstrating that multiple fqs can be used here...
fobj = run(x = "fastq_bam_bwa", samplename = "samp1", execute = TRUE,
		fqs1 = rep(paste0(fqpath, "/brca.example.illumina.0.1.fastq"), 2), 
		fqs2 = rep(paste0(fqpath, "/brca.example.illumina.0.2.fastq"), 2))
```

OR from R using:



Refer to the help pages for more details on the [run function](http://docs.flowr.space/rd.html#run).


#### Details regarding cluster submission

The **run** function performs several steps, finally submitting the commands to the cluster. 
It may be useful to go through these steps to understand the details.


**1. Get user inputs**

Using the name of the pipeline, `run` fetches it in various places inclusing `~/flowr/pipelines`.


```{r message=FALSE, warning=FALSE, results='hide'}
library(flowr)
setwd("~/flowr/pipelines")
source("fastq_bam_bwa.R")
## this may throw a warning if paths do not exist
## if you have used modules instead of full paths please ignore the warnings
load_opts("fastq_bam_bwa.conf") 

## Get example input
## these can be a vector of multiple paired-end files
## OR multiple single-end files
fqs1 = "~/flowr/genomes/target_brca_example/brca.example.illumina.0.1.fastq"
fqs2 = "~/flowr/genomes/target_brca_example/brca.example.illumina.0.2.fastq"
samp = "samplename"

## optionally specify the center, lane, platform etc.
set_opts(rg_center = "the_institute", rg_lane = "1")

## **Note:** load_opts checks if variables ending with 
## _exe, _path, _dir etc. exist or not.
## make sure they are all correct. 
## Ignore the warnings, if instead of specifying full path to a tool 
## you are using the module command.
```

Refer to the help pages of [fetch_pipes](http://docs.flowr.space/rd.html#fetch_pipes) and 
[fetch_pipes](http://docs.flowr.space/rd.html#load_opts) for more details.


**2. Read flow definition**


```{r }
def = as.flowdef("fastq_bam_bwa.def")
```

The plot would work only if you have X11 etc enabled, i.e. if you logged into the cluster using `ssh -X` (or `ssh -Y`).

Optionally, one can edit all config files on their own machine, debug and sort issues; when done, upload them to the cluster and submit.

```{r}
plot_flow(def) ## on a cluster, only works if graphics X11 is enabled. ssh -X
```

**3. Create a table with all commands to run**

We use the function `fastq_bam_bwa` to create a [flow mat](http://docs.flowr.space/docs.html#flow_mat).

```{r message=FALSE}
## run the module and create a flow mat, with all the commands 
out = fastq_bam_bwa(fqs1, fqs2, samplename = samp)

## optionally, write this to a file (a simple tab delimited table)
write_sheet(out$flowmat, "fastq_bam_bwa.tsv")
```


**4. Executing on the computing cluster**


Now we can submit this to the cluster using:

```{r eval=FALSE}
fobj2 = to_flow(x='~/flowr/pipelines/fastq_bam_bwa.tsv', 
                def='~/flowr/pipelines/fastq_bam_bwa.def',
                name = "fastq_bam_bwa", 
                execute=TRUE)
```

OR from the terminal using:

```
flowmat=~/flowr/pipelines/fastq_bam_bwa.tsv
flowdef=~/flowr/pipelines/fastq_bam_bwa.def
flowr to_flow x=$flowmat def=$flowdef name=fastq_bam_bwa execute=TRUE
```

<div class="alert alert-info" role="alert">
**Tip**:
This example shows a single sample, but you may have
as many samples in the flowmat. 
In case of multiple samples, the **samplename** column is used to group commands 
and each set if submitted as a individual flow.
</div>


**Several other functions, one may use after submission:**

*checking the status:*

```
## from R:
status(x="~/flowr/runs/fastq_bam_bwa*")

## OR from terminal using:
flowr status x=~/flowr/runs/fastq_bam_bwa*
```
```
|          | total| started| completed| exit_status|status     |
|:---------|-----:|-------:|---------:|-----------:|:----------|
|001.aln1  |     1|       1|         0|           0|processing |
|002.aln2  |     1|       1|         0|           0|processing |
|003.sampe |     1|       0|         0|           0|pending    |
|004.fixrg |     1|       0|         0|           0|pending    |
|005.merge |     1|       0|         0|           0|pending    |
```


Additionally, you may **[kill](http://docs.flowr.space/rd.html#kill)** or 
**[rerun](http://docs.flowr.space/rd.html#kill)** the flow as well. 


```
flowr kill x=~/flowr/runs/fastq_bam_bwa*
flowr rerun x=~/flowr/runs/<full path of the flow> start_from=fixrg
```

Please use the respective help pages for more details.

<br>
<br>
<br>


