---
title: "flowr"
subtitle: "Streamlining Workflows"
author: Sahil Seth
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document: 
    keep_md: yes
    toc: true
vignette: >
    %\VignetteIndexEntry{Tutorial on Building Pipelines}
    %\VignetteEngine{knitr::rmarkdown}
    \usepackage[utf8]{inputenc}
packagedocs:
    toc: true
    toc_depth: 4
navpills: |
  <li><a href='docs.html'>Overview</a></li>
  <li><a href='install.html'>Install</a></li>
  <li class="active"><a href='Tutorial.html'>Tutorial</a></li>
  <li><a href='rd.html'>Help</a></li>
  <li><a href='news.html'>News</a></li>
  <li><a href='https://github.com/sahilseth/flowr'>Github <i class='fa fa-github'></i></a></li>
---

```{r libs_tut, echo = FALSE, message = FALSE}
library(knitr)
library(flowr)
```

# Tutorial: building a pipeline

```{r example1, cache = FALSE, echo=FALSE}
read_chunk(system.file('pipelines', 'sleep_pipe.R', package = 'flowr'))
```



A pipeline consists of several pieces, most essential of which is a function which generates a [flowmat](http://docs.flowr.space/docs.html#flow_mat).
Additionally, we need a [flow definition](http://docs.flowr.space/docs.html#flow_definition), which descibes flow of the pipeline. 
These three files are available under the [pipelines](https://github.com/sahilseth/flowr/tree/master/inst/pipelines) folder on github.

```
## to follow this tutorial, you may download them:
url=https://raw.githubusercontent.com/sahilseth/flowr/master/inst/pipelines
cd ~/flowr/pipelines
wget $url/sleep_pipe.R             ## A R script, with sleep_pipe(), which creates a flowmat
wget $url/sleep_pipe.def           ## A tab-delimited flow definition file
wget $url/sleep_pipe.conf          ## An *optional* tab-delim conf file, defining default params
```



To run the aforementioned pipeline, we would follow through these steps:

```{r eval=FALSE}
## Single step submission:
fobj = run("sleep_pipe", execute = TRUE); 

## Details of the above step:
setwd("~/flowr/pipelines")
## behind the scenes, run does the following:
## optionally, load default parameters
load_opts("sleep_pipe.conf") 

## get sleep_pipe() function
source("sleep_pipe.R") 

## create a flowmat
flowmat = sleep_pipe()

## read a flow definition.
flowdef = as.flowdef("sleep_pipe.def")

## create flow and submit to cluster
fobj = to_flow(flowmat, flowdef, execute = TRUE)
```




## Creating Modules/Pipelines


<div class="alert alert-info" role="alert">
**module:** A R function which creates a flow mat, is a module. Using **module + flowdef**, we can run a pipeline.
</div>


The `sleep_pipe` calls the three other functions (**modules**); fetches flowmat from each, then rbinds them,
creating a larger flowmat. You may refer to the [sleep_pipe.R](https://github.com/sahilseth/flowr/blob/master/inst/pipelines/sleep_pipe.R)
file for the source.

```{r define_modules, echo=FALSE}

```

```{r define_pipeline}

```


```{r}
## create a flow matrix
out = sleep_pipe(x = 3, "sample1")
flowmat = out$flowmat
```

```{r, echo=FALSE}
kable(flowmat)
```


Next, we need a flow definition.

## Creating Flow Definition

flowr enables us to quickly create a skeleton flow definition using a flowmat, which we can then alter to suit our needs. A handy function
to_flowdef, accepts a `flowmat` and creates a flow definition. 

<div class="alert alert-info" role="alert">
The default skeleton takes a very conservative approach, creating all submissions as `serial` and all dependencies as `gather`. This ensures robustness, compromising efficiency. 
</div>

```{r plot_skeleton_def, message=FALSE}
def = to_flowdef(flowmat) ## create a skeleton flow definition
suppressMessages(plot_flow(def))
```

We can make the following changes to make this more efficient (run steps in parallel):


- multiple sleep commands would run as `scatter`/parallel (`none`)
- For each sleep, create_tmp creates a tmp file (`serial`)
- All tmp files are merged; when all are complete (`gather`)
- Then we get size on the resulting file (`serial`)

*dependencies mentioned in ()*


```{r message=FALSE}
def$sub_type = c("scatter", "scatter", "serial", "serial")
def$dep_type = c("none", "serial", "gather", "serial")
kable(def)
```

```{r plot_tweaked_def, message=FALSE, echo = FALSE}
suppressMessages(plot_flow(def))
```


<div class="alert alert-info" role="alert">
**Tip:** Alternatively, one may write this to a file 
(`write_sheet(def, "sleep_pipe.def")`), make changes in a text editor and read it again (`as.flowdef("sleep_pipe.def")`.
</div>

## Create flow, submit to cluster

**Next, we create a flow object:**

```{r, message=FALSE}
fobj = to_flow(flowmat, def, flowname = "sleep_pipe")
```

**Finally, we can submit this to the cluster:**
```{r eval=FALSE}
plot_flow(fobj)
submit_flow(fobj) ## dry run
fobj2 = submit_flow(fobj, execute = TRUE) ## submission to LSF cluster

## after submission, we can use the following:
status(fobj2) ## check status
rerun(fobj2)  ## re-run from a intermediate step
kill(fobj2)   ## kill it!
```


<!----


We then define another function `sleep_pipe` which calls the above defined **modules**; fetches flowmat from each, 
creating a larger flowmat. This time we will define a flowdef for the `sleep_pipe` function, elevating its status from
module to a pipeline.


This time we will define a flowdef for the `sleep_pipe` function, elevating its status from
module to a pipeline.




Here are a few examples of modules, three functions `sleep`, `create_tmp` and `merge_size` each returning a flowmat.

We believe pipeline and modules may be interchangeble, in the sense that a *smaller* pipeline may be 
included as part of a larger pipeline.
In flowr a module OR pipeline always returns a flowmat.
The only difference being, a pipeline also has a correspomding flow definition file. 


<div class="alert alert-info" role="alert">
As such, creating a flow definition for a module enables flowr
to run it, hence a module **elevates**, becoming a pipeline.
This lets the user mix and match several modules/pipelines to create a customized larger pipeline(s).
</div>
-->


<script>
  (function(i,s,o,g,r,a,m){i['GoogleAnalyticsObject']=r;i[r]=i[r]||function(){
  (i[r].q=i[r].q||[]).push(arguments)},i[r].l=1*new Date();a=s.createElement(o),
  m=s.getElementsByTagName(o)[0];a.async=1;a.src=g;m.parentNode.insertBefore(a,m)
  })(window,document,'script','//www.google-analytics.com/analytics.js','ga');

  ga('create', 'UA-68378910-1', 'auto');
  ga('send', 'pageview');

</script>
