---
title: "flowr simple examples"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
vignette: >
  %\VignetteIndexEntry{flowr simple example}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, echo = FALSE, message = FALSE}
library(knitr)
knitr::opts_chunk$set(
  comment = "#>",
  error = FALSE,
  tidy = FALSE,
	fig.cap = ""
)
library(flowr)

```


As of now we have tested this on the following clusters:

```{r echo=FALSE}
exdata = file.path(system.file(package = "flowr"), "misc")
read_sheet(file.path(exdata, "platforms_supported.txt"))


```

\*queue short-name used in `flow <https://github.com/sahilseth/flow>`__

There are several `job
scheduling <http://en.wikipedia.org/wiki/Job_scheduler>`__ systems
available and we try to support the major players. Adding support is
quite easy if we have access to them. Your favourite not in the list?
Send a `message <mailto:sahil.seth@me.com>`__

-  `PBS wiki <http://en.wikipedia.org/wiki/Portable_Batch_System>`__:

   -  `Torque
      wiki <http://en.wikipedia.org/wiki/TORQUE_Resource_Manager>`__
   -  MD Anderson, `University of
      Houston <http://www.rcc.uh.edu/hpc-docs/49-using-torque-to-submit-and-monitor-jobs.html>`__

-  `LSF <http://en.wikipedia.org/wiki/Platform_LSF>`__:

   -  Harvard Medicla School uses: `LSF HPC
      7 <https://wiki.med.harvard.edu/Orchestra/IntroductionToLSF>`__.
   -  Also Used at
      `Broad <https://www.broadinstitute.org/gatk/guide/article?id=1311>`__

-  `SGE <http://en.wikipedia.org/wiki/Sun_Grid_Engine>`__
  -  A tutorial for `Sun Grid
      Engine <https://sites.google.com/site/anshulkundaje/inotes/programming/clustersubmit/sun-grid-engine>`__,
      you may also see another at
      `JHSPH <http://www.biostat.jhsph.edu/bit/cluster-usage.html>`__.
      Dependecy info
      `here <https://wiki.duke.edu/display/SCSC/SGE+Job+Dependencies>`__

http://en.wikipedia.org/wiki/Comparison_of_cluster_software

