---
title: "Quick Start Example"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quick Start Example}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

Get me started
-------------


```{r eval=FALSE}
install.packages('devtools')
devtools::install_github("sahilseth/flowr")
```

Run a setup function which copies 'flowr' Rscript to subsetquent steps easier.
More on this [here](https://github.com/sahilseth/rfun).

```{r}
library(flowr)
setup()
```


# Create a flow using example data
```{r}
exdata = file.path(system.file(package = "flowr"), "extdata")
flow_mat = read_sheet(file.path(exdata, "example1_flow_mat.txt"))
flow_def = read_sheet(file.path(exdata, "example1_flow_def.txt"))

flow_mat = subset(flow_mat, samplename == "sample1")

fobj <- to_flow(x = flow_mat, def = flow_def, 
	flowname = "example1",
	platform = "lsf")

```

# Ingredient 1: Commands to run (flow_mat)
```{r}
kable(head(flow_mat))
```

# Ingredient 2: Flow Definition (flow_def)
```{r}
kable(flow_def)
```

**The above table basically translates to**:

- `sleep`: Run all 10 sleep jobs for given sample
- `tmp`: Create 10 temporary files, after sleep jobs are complete
	- dependency is serial, tmp jobs does not wait for all sleep jobs to complete. 
	- This is a one-to-one relationship
- `merge`: When all `tmp` are complete, merge them
- `size`: get their size when merge is complete

# Plot describing the definition
```{r plot_example1, fig.cap="Flow chart describing process for example 1", fig.height=10, fig.width=8}
plot_flow(fobj)
```


# Dry run (submit)
```{r eval=FALSE}
submit_flow(fobj)
```

```
Test Successful!
You may check this folder for consistency. Also you may re-run submit with execute=TRUE
 ~/flowr/type1-20150520-15-18-27-5mSd32G0
```

# Submit to the cluster
```{r eval=FALSE}
submit_flow(fobj, execute = TRUE)
```

```
Flow has been submitted. Track it from terminal using:
flowr::status(x="~/flowr/type1-20150520-15-18-46-sySOzZnE")
OR
flowr status x=~/flowr/type1-20150520-15-18-46-sySOzZnE
```


# Check the status

```
flowr status x=~/flowr/type1-20150520-15-18-46-sySOzZnE
```

```
Loading required package: shape
Flowr: streamlining workflows
Showing status of: /rsrch2/iacs/iacs_dep/sseth/flowr/type1-20150520-15-18-46-sySOzZnE


|          | total| started| completed| exit_status|
|:---------|-----:|-------:|---------:|-----------:|
|001.sleep |    10|      10|        10|           0|
|002.tmp   |    10|      10|        10|           0|
|003.merge |     1|       1|         1|           0|
|004.size  |     1|       1|         1|           0|
```
