---
title: "Quick Start Example"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quick Start Example}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---


```{r eval=FALSE}
install.packages('devtools')
devtools::install_github("sahilseth/flow")
```

Run a setup function which copies 'flowr' Rscript to subsetquent steps easier.
More on this [here](https://github.com/sahilseth/rfun).

```{r}
library(flowr)
setup()
```


# Create a flow using example data
```{r}
exdata = file.path(system.file(package = "flowr"), "extdata")
flow_mat = read_sheet(file.path(exdata, "example1_flow_mat.txt"))
flow_def = read_sheet(file.path(exdata, "example1_flow_def.txt"))

flow_mat = subset(flow_mat, samplename == "sample1")

fobj <- to_flow(x = flow_mat, def = flow_def, 
	flowname = "example1",
	platform = "lsf")

```


```{r}
plot_flow(fobj)
```

## For Figure1 the following process would be followed:

- `sleep`: Run all 10 sleep jobs for given sample
- `tmp`: Create 10 temporary files, after sleep jobs are complete
	- dependency is serial, tmp jobs does not wait for all sleep jobs to complete. 
	- This is a one-to-one relationship
- `merge`: When all `tmp` are complete, merge them
- `size`: get their size when merge is complete

```{r eval=FALSE}
submit_flow(fobj)
submit_flow(fobj, execute = TRUE)
```

