---
title: "flowr simple examples"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{flowr simple example}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---


```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  error = FALSE,
  tidy = FALSE
)
library(flowr)
```

### Assuming we are working on a LSF platform, we define it using type. Currently we support LSF and Torque (SGE). 
```{r getqobj}
qobj <- queue(type = "lsf", queue = "normal")
```

### creating a pipeline definition: Style1
- Each module here has a single job
- job2 and job3 would start after 1 is complete
```{r plot_simpleflow}
job1 <- job(name = "myjob1", cmds = "sleep1", q_obj = qobj)
job2 <- job(name = "myjob2", cmds = "sleep2", q_obj = qobj, previous_job = "myjob1", dependency_type = "serial")
job3 <- job(name = "myjob3", cmds = "sleep3", q_obj = qobj, previous_job = "myjob1", dependency_type = "serial")
fobj <- flow(name = "myflow", jobs = list(job1, job2, job3), desc="description")
#source('/DRIVES/Dropbox2/Dropbox/public/github_flow/R/plot-funcs.R')
plot_flow(fobj)
```

```{r}
dat <- flowr:::.create_jobs_mat(fobj)
knitr:::kable(dat)
```

## An example of Multiple jobs

### Submission types
- scatter: submit all commands as parallel independent jobs
- serial: run these commands sequentuially one after the other

### Dependency types
- none: independent job
- serial: *one to one* relationship with previous job
- gather: *many to one* wait for **all** commands in previous job to finish then start current


### Serial: one to one relationship
- All commads in 'job1' are submitted, and those is 'jobs2' *wait* for those in 'job1' to complete.
- Commands in 'job2' are serially dependent on 'job1'
- Both jobs are submitted as parallel (*scatter*), i.e. there is not **intra** dependency.
- so previous job submission: `scatter`, and current job's dependency type `serial`
```{r}
cmds = rep("sleep 5", 10)
jobj1 <- job(q_obj=qobj, cmd = cmds, submission_type = "scatter", name = "job1")
jobj2 <- job(q_obj=qobj, name = "job2", cmd = cmds, submission_type = "scatter", 
             dependency_type = "serial", previous_job = "job1")
fobj <- flow(jobs = list(jobj1, jobj2))
plot_flow(fobj)
```

### Gather: many to one relationship
- makes sense when previous job had many commands running in parallel and current job would wait for all
- so previous job submission: `scatter`, and current job's dependency type `gather`
```{r}
jobj1 <- job(q_obj=qobj, cmd = cmds, submission_type = "scatter", name = "job1")
jobj2 <- job(q_obj=qobj, name = "job2", cmd = cmds, submission_type = "scatter", 
             dependency_type = "gather", previous_job = "job1")
fobj <- flow(jobs = list(jobj1, jobj2))
plot_flow(fobj)
```

### Burst: one to many relationship
- makes sense when previous job had one command current job would split and submit several jobs in parallel
- so previous job submission_type: `serial`, and current job's dependency type `burst`, with a submission type: `scatter`
```{r}
jobj1 <- job(q_obj=qobj, cmd = cmds, submission_type = "serial", name = "job1")
jobj2 <- job(q_obj=qobj, name = "job2", cmd = cmds, submission_type = "scatter", 
             dependency_type = "burst", previous_job = "job1")
fobj <- flow(jobs = list(jobj1, jobj2))
plot_flow(fobj)
```


For more details look at [Building flowr recepies](buid_main.html)
